---
title: "Real State Exploration"
author: "Gabriel Bessa"
date: "August 17, 2018"
output:
  html_document:
    df_print: paged
---

# Introduction

Moving away from home for the first time is often a experience that most of people will always remember. This experience can be either really good or really bad. I know that someday it will be my turn, so I decided to prepare my self to do it.

This report will explore a dataset containing rental prices and attributes for approximately 2218 apartments that are located on my hometown (Belo Horizonte, Brazil). This data set was crawled from [NetImoveis](https://www.netimoveis.com/) using the [NetSpider](https://github.com/gabrielvcbessa/real-state-crawlers/blob/master/net_spider.py) crawler.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)

df.raw <- read.csv('net_apartments.csv')
df.raw[df.raw == ''] <- NA # We are going to threat spaces as NA's

df <- df.raw[, colSums(is.na(df.raw)) < nrow(df.raw)] %>% 
  select(vagasGaragem,
         areaRealPrivativa,
         sistemaDeAlarme,
         salaoDeJogos,
         mobiliado,
         guarita,
         gasCanalizado,
         valorImovel,
         playGround,
         piscina,
         churrasqueira,
         pontuacao,
         latitude2,
         longitude2,
         quartos,
         valorCondominio,
         valorLocacao,
         tipoImovel1Al,
         sauna,
         academia,
         nomeBairroAl,
         imovelSan_Id,
         suites,
         portariaPermanente) %>% 
  rename(id = imovelSan_Id,
         size = areaRealPrivativa,
         propertyType = tipoImovel1Al,
         rentalPrice = valorLocacao,
         propertyPrice = valorImovel,
         parkingSpaces = vagasGaragem,
         alarmSystem = sistemaDeAlarme,
         gameRoom = salaoDeJogos,
         furnished = mobiliado,
         securityCabin = guarita,
         pipedGas = gasCanalizado,
         swimmingPool = piscina,
         barbecueGrill = churrasqueira,
         score = pontuacao,
         latitude = latitude2,
         longitude = longitude2,
         bedrooms = quartos,
         monthlyFees = valorCondominio,
         steamRoom = sauna,
         gym = academia,
         neighborhood = nomeBairroAl,
         permanentDoorman = portariaPermanente) %>% 
  mutate(m2RentalPrice = rentalPrice / size)

df <- transform(df, link = paste('https://www.netimoveis.com/imovel/dummy', id, sep = '/'))
```

# Analysis
## Plotting locations

Usually we starting looking for apartments on certain locations that we have interest in, ignoring others that maybe aren't that close to our interest point, but can be a good option if you consider the price. So, remove this bias, we want to have a dataset that contains apartments from different regions. Let's see how our apartments are distributed:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggmap)
library(maps)

bbox.apartments <- make_bbox(df$longitude, df$latitude, f = 0.05)
map.apartments <- get_map(location = bbox.apartments,
                          source = 'stamen')

ggmap(map.apartments) +
  geom_point(aes(x = longitude, y = latitude), 
             data = df, alpha = .75) +
  theme(plot.margin=grid::unit(c(0, 0, 0, 0), 'mm'))

```

As we can see, we have some misplaced points. Few of them have 0 value for the latitude and longitude, and others are placed on different brazilian capitals. We are going to impute those values and generate our map again. Let's see how our new plot looks like:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gridExtra)

outlier_values <- boxplot.stats(df$longitude)$out
outlier_indexes <- which(df$longitude %in% outlier_values)
outliers <- df[outlier_indexes, ]

# We are going to input some values to the missing locations
df.by_neigh <- df[-outlier_indexes, ] %>% 
  group_by(neighborhood) %>% 
  summarise(latitude = mean(latitude),
            longitude = mean(longitude))

outliers.imputed <- 
  transform(outliers,
            latitude = df.by_neigh[neighborhood == neighborhood, ]$latitude,
            longitude = df.by_neigh[neighborhood == neighborhood, ]$longitude)

df.imputed <- df
df.imputed[outlier_indexes, ] <- outliers.imputed

bbox.imputed <- make_bbox(df.imputed$longitude, df.imputed$latitude, f = 0.1)
map.imputed <- get_map(location = bbox.imputed,
                          source = 'stamen')

# Plotting all apartments
p1 <- ggmap(map.imputed) +
  geom_point(aes(x = longitude, y = latitude), 
             data = df.imputed, alpha = .25, size = .75) +
  labs(title = 'All apartments')

df.imputed.by_neigh <- df.imputed %>% 
  group_by(neighborhood) %>% 
  summarise(count = n(),
            longitude = mean(longitude),
            latitude = mean(latitude))

bbox.imputed.by_neigh <- 
  make_bbox(df.imputed.by_neigh$longitude, df.imputed.by_neigh$latitude, f = 0.1)

map.imputed.by_neigh <- get_map(location = bbox.imputed.by_neigh,
                                source = 'stamen')

# Plotting apartments by neighborhood
p2 <- ggmap(map.imputed, padding = 0) +
  geom_point(aes(x = longitude, y = latitude, size = count), 
             data = df.imputed.by_neigh, alpha = .55) +
  labs(title = 'Apartments by neighborhood')

grid.arrange(p1, p2, ncol = 2)
```

